---
title: "Aids Drug"
author: "Eduardo Noriega"
date: "9/28/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)

females=filter(triUmp,gender== "F")
IVDrugUsers=filter(triUmp,DrugUser=="yes")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aids drug (triUmpsal) is under test in clinical trials. Consider the DrugUsers, does this drug seem effective on this group?  Hint - use dplyr to obtain that sub-population with the filter command.  Next, repeat the analysis on just females, does this drug seem effective on women?

<!-- Be sure to produce Descriptive Results and Inferential Results for both groups. !-->

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed, for example sub-setting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

1. Is the vaccine effective for drug users? 

$H_0$: The vaccine is not effective for drug users.  
$H_A$: The vaccine is effective for drug users.

2. Is the the vaccine effective for women?

$H_0$: The vaccine is not effective for females.  
$H_A$: The vaccine is effective for females.  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe patterns, if any, that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming to or departing from what is expected by the null hypothesis !-->

For Females

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

The graph shows that 60% of females with the vaccine are likely to get aids, versus the ones that received the placebo which was only 40% of females.

Drug Users  

```{r}
barchartGC(~infected + treatment,data=IVDrugUsers)
barchartGC(~infected + treatment,data=IVDrugUsers, type="percent")

```

The vaccines shows to be effective with the drug users, which about 25% of them became infected with aids, versus those who received the placebo which is close to 70%.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns, if any, that exist in the data as focused toward the research question!-->

Females

```{r}
tableF <- xtabs(~infected + treatment, data=females)
rowPerc(tableF)
colPerc(tableF)
```
This chart shows that the females that received the placebo which were
41.18%, got infected, compared to 58.82% who did receive the vaccine. Females who received the vaccine are more susceptible to contracting aids.

Drug Users
```{r}
tableIV<- xtabs(~infected + treatment, data=IVDrugUsers)
rowPerc(tableIV)
colPerc(tableIV)
```
Among drug users, those who received the placebo were 75.71% of them contracted aids, compared to those who received the vaccine,24.29% got infected with aids. So far this is showing that the vaccine is more effective towards those who are drug users.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->

Females
```{r}
chisq.test(tableF)
chisqtestGC(tableF)
fisher.test(tableF)
```



IV Drug Users
```{r}
chisq.test(tableIV)
chisqtestGC(tableIV)
fisher.test(tableIV)
```


  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!

Report on both Drug Users and Females.  Be sure to include the odds for both groups.
-->
 The is significant difference between drug users and women for those who get the vaccine and contracted AIDS. 

1. The vaccine is more effective for the drug users, showing that they are less likely to get infected with AIDS.The null hypothesis is rejected due to the p-value being less than 0.05, which is 0.0000000423 as shown in the fisher test.

2. Women who received the vaccine did have an increase is getting infected with AIDS, versus women who received the placebo. Per the chi square test the p-value is 0, therefore the null hypothesis is rejected as well.
 
